---
title: "R Basics"
format: pdf
editor: visual
---

## Declaring Variables

In R, you assign values to variables using the `<-` operator. This is how you store numbers, text, or logical values (TRUE/FALSE).

```{r}
x <- 10           # numeric
string <- "CPAL"   # character
bool <- TRUE      # logical
```

You can also use `=` to assign values, but `<-` is the preferred style in R.

## Looking Up Functions

Use `?function_name` or `help()` to look up how a function works. Use `??keyword` to search for functions or help topics related to a word.

```{r}
?mean      # Opens help for the mean() function
help(sum)  # Opens help for sum()
??paste # Opens topics related to regression
```

## Basic Data Types

R has several basic data types. You can use `class()` or `str()` to check them.

```{r}
class(5)
str(5)
class("text")
class(TRUE)
class(c(1, 2, 3))
```

## Reading/Writing Data

Use `read.csv()` to read in data provided in a .csv. Use `write.csv()` to create a .csv from your dataframe.

```{r}
wic <- read.csv("data/WIC Eligibility Estimates by County and Year.csv")
write.csv(wic, 'data/WIC Eligibility Estimates Output.csv') 
```

For Excel sheets, use `read_xlsx()` from `library(readxl)` to read in data. Use `write_xlsx()` to create a .xlsx from your dataframe.

```{r}
library(readxl)
dallasCollege <- read_xlsx('data/Housing_Data_for_Export_v3 (1).xlsx')
```

For Google Sheets, use `read_sheet()` from the library `googlesheets4` to read in data. Use `write_sheet()` to write data to a Google Sheet
